---
title: "Introduction to reticulate"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R interface to Python}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Overview

The **reticulate** package provides an R interface to Python modules, classes, and functions. For example, this code imports the Python `os` module and calls some functions within it:

```{r}
library(reticulate)
os <- import("os")
os$chdir("tests")
os$getcwd()
```

Functions and other data within Python modules and classes can be accessed via the `$` operator (analogous to the way you would interact with an R list, environment, or reference class).

The **reticulate** package is compatible with all versions of Python >= 2.7. Integration with NumPy is optional and requires NumPy >= 1.6.

## Type Conversions

When calling into Python, R data types are automatically converted to their equivalent Python types. When values are returned from Python to R they are converted back to R types. Types are converted as follows:

| R  | Python | Examples |
|---|---|---|
| Single-element vector   | Scalar |  `1`, `1L`, `TRUE`, `"foo"` |
| Multi-element vector | List  |  `c(1.0, 2.0, 3.0)`, `c(1L, 2L, 3L)` |
| List of multiple types  | Tuple  |  `list(1L, TRUE, "foo")`
| Named list | Dict  |  `list(a = 1L, b = 2.0)`, `dict(x = x_data)`
| Matrix/Array | NumPy ndarray  | `matrix(c(1,2,3,4), nrow = 2, ncol = 2)`
| Function | Python function | `function(x) x + 1`
| NULL, TRUE, FALSE  | None, True, False  |  `NULL`, `TRUE`, `FALSE`

If a Python object of a custom class is returned then an R reference to that object is returned. You can call methods and access properties of the object just as if it was an instance of an R reference class.

## Importing Modules

The `import` function can be used to import any Python module. For example:

```{r}
difflib <- import("difflib")
difflib$ndiff(foo, bar)

filecmp <- import("filecmp")
filecmp$cmp(dir1, dir2)
```

The `import_main()` and `import_builtins()` functions give you access to the main module where code is executed by default and the collection of built in Python functions. For example:

```{r}
main <- import_main()

py <- import_builtins()
py$print('foo')
```

The main module is generally useful if you have executed Python code from a file or string and want to get access to it's results (see the section below for more details).

### Object Conversion

By default when Python objects are returned to R they are converted to their equivalent R types. However, if you'd rather make conversion from Python to R explicit and deal in native Python objects by default you can pass `convert = FALSE` to the `import` function. In this case Python to R conversion will be disabled for the module returned from `import`. For example:

```{r}
# import numpy and specify no automatic Python to R conversion
np <- import("numpy", convert = FALSE)

# do some array manipulations with NumPy
a <- np$array(c(1:4))
sum <- a$cumsum()

# convert to R explicitly at the end
py_to_r(sum)
```

As illustrated above, if you need access to an R object at end of your computations you can call the `py_to_r()` function explicitly.

## Executing Code

You can execute Python code within the main module using the `py_run_file` and `py_run_string` functions. These functions both return a reference to the main Python module so you can access the results of their execution. For example:

```{r}
py_run_file("script.py")

main <- py_run_string("x = 10")
main$x
```

## Getting Help

You can print documentation on any Python object using the `py_help()` function. For example:

```r
os <- import("os")
py_help(os$chdir)
```

## Lists, Tuples, and Dictionaries

The automatic conversion of R types to Python types works well in most cases, but occasionally you will need to be more explicit on the R side to provide Python the type it expects.

For example, if a Python API requires a list and you pass a single element R vector it will be converted to a Python scalar. To overcome this simply use the R `list` function explicitly:

```r
foo$bar(indexes = list(42L))
```

Similarly, a Python API might require a `tuple` rather than a list. In that case you can use the `tuple()` function:

```r
tuple("a", "b", "c")
```

R named lists are converted to Python dictionaries however you can also explicitly create a Python dictionary using the `dict()` function:

```r
dict(foo = "bar", index = 42L)
```

This might be useful if you need to pass a dictionary that uses a more complex object (as opposed to a string) as it's key.


## With Contexts

The R `with` generic function can be used to interact with Python context manager objects (in Python you use the `with` keyword to do the same). For example:

```{r}
py <- import_builtins()
with(py$open("output.txt", "w") %as% file, {
  file$write("Hello, there!")
})
```

This example opens a file and ensures that it is automatically closed at the end of the with block. Note the use of the `%as%` operator to alias the object created by the context manager.

## Iterators

If a Python API returns an [iterator or generator](http://anandology.com/python-practice-book/iterators.html) you can interact with it using the `iterate()` function. The `iterate()` function can be used to apply an R function to each item yielded by the iterator:

```{r}
iterate(iter, print)
```

If you don't pass a function to `iterate` the results will be collected into an R vector:

```{r}
results <- iterate(iter)
```

Note that the `Iterators` will be drained of their values by `iterate()`:

```{r}
a <- iterate(iter) # results are not empty
b <- iterate(iter) # results are empty since items have already been drained
```

### Element Level Iteration

You can also iterate on an element-by-element basis using the `iter_next()` function. For example:

```{r}
while (TRUE) {
  item <- iter_next(iter)
  if (is.na(item))
    break
}
```

By default `iter_next()` will return `NA` when the iteration is complete but you can provide a custom value and it will be returned instead. For example:

```{r}
while (TRUE) {
  item <- iter_next(iter, completed = NULL)
  if (is.null(item))
    break
}
```

Note that some iterators/genrators in Python are infinite. In that case the caller will need custom logic to determine when to terminate the loop.


## Advanced Functions

There are several more advanced functions available that are useful principally when creating high level R interfaces for Python libraries.

### Python Objects

Typically interacting with Python objects from R involves using the `$` operator to access whatever properties for functions of the object you need. When using the `$`, Python objects are automatically converted to their R equivalents when possible. The following functions enable you to interact with Python objects at a lower level (e.g. no conversion to R is done unless you explicitly call the `py_to_r` function):

| Function  | Description | 
|----------------|------------------------------------------------------------------|
| `py_has_attr()` | Check if an object has a specified attribute.  | 
| `py_get_attr()` | Get an attribute of a Python object.  | 
| `py_set_attr()` | Set an attribute of a Python object.  | 
| `py_list_attributes()` | List all attributes of a Python object. |
| `py_call()` | Call a Python callable object with the specified arguments.  | 
| `py_to_r()` | Convert a Python object to it's R equivalent |
| `r_to_py()` | Convert an R object to it's Python equivalent |

### Configuration

The following functions enable you to query for information about the Python configuration available on the current system.

| Function  | Description | 
|----------------|------------------------------------------------------------------|
| `py_available()` | Check whether a Python interface is available on this system. |
| `py_numpy_available()` | Check whether the R interface to NumPy is available (requires NumPy >= 1.6) |
| `py_module_available()` | Check whether a Python module is available on this system. |
| `py_config()` | Get information on the location and version of Python in use.  | 


### Output Control

These functions enable you to capture or suppress output from Python:

| Function  | Description | 
|----------------|------------------------------------------------------------------|
| `py_capture_output()` | Capture Python output for the specified expression and return it as an R character vector. | 
| `py_suppress_warnings()` | Execute the specified expression, suppressing the display Python warnings.   | 


### Miscellaneous

The functions provide miscellaneous other lower-level capabilities:

| Function  | Description | 
|----------------|------------------------------------------------------------------|
| `py_unicode()` | Convert a string to a Python unicode object. |
| `py_str()` | Get the string representation of Python object. |
| `py_is_null_xptr()` | Check whether a Python object is a null externalptr. |
| `py_validate_xptr()` | Check whether a Python object is a null externalptr and throw an error if it is. |


## Learning More

The following articles cover additional aspects of using **reticulate**:

- [Python Version Configuration](versions.html)

- [Using reticulate in an R Package](package.html)

- [Arrays in R and Python](arrays.html)


