---
title: "R Interface to Python"
output: 
  rmarkdown::html_vignette:
    readme: true
vignette: >
  %\VignetteIndexEntry{R interface to Python}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

<div id="author-region" style="display:none; margin-bottom: 15px;">
J.J. Allaire &mdash; `r Sys.Date()`
</div>

```{js, eval=TRUE}
document.getElementById("author-region").style.display = 'inherit';
```

## Overview

The **reticulate** package provides an R interface to Python modules, classes, and functions. For example, this code imports the Python `os` module and calls some functions within it:

```{r}
library(reticulate)
os <- import("os")
os$chdir("tests")
os$getcwd()
```

Functions and other data within Python modules and classes can be accessed via the `$` operator (analogous to the way you would interact with an R list, environment, or reference class).

When calling into Python R data types are automatically converted to their equivalent Python types. When values are returned from Python to R they are converted back to R types. Types are converted as follows:

| R  | Python | Examples |
|---|---|---|
| Single-element vector   | Scalar |  `1`, `1L`, `TRUE`, `"foo"` |
| Multi-element vector | List  |  `c(1.0, 2.0, 3.0)`, `c(1L, 2L, 3L)` |
| List of multiple types  | Tuple  |  `list(1L, TRUE, "foo")`
| Named list | Dict  |  `list(a = 1L, b = 2.0)`, `dict(x = x_data)`
| Matrix/Array | NumPy ndarray  | `matrix(c(1,2,3,4), nrow = 2, ncol = 2)`
| Function | Python function | `function(x) x + 1`
| NULL, TRUE, FALSE  | None, True, False  |  `NULL`, `TRUE`, `FALSE`

If a Python object of a custom class is returned then an R reference to that object is returned. You can call methods and access properties of the object just as if it was an instance of an R reference class.

The **reticulate** package is compatible with all versions of Python >= 2.7 and in addition requires NumPy >= 1.11.

## Installation

You can install from GitHub as follows:

```{r}
devtools::install_github("rstudio/reticulate")
```

Note that the includes native C/C++ code so it's installation requires [R Tools](https://cran.r-project.org/bin/windows/Rtools/) on Windows and [Command Line Tools](http://osxdaily.com/2014/02/12/install-command-line-tools-mac-os-x/) on OS X. If the package installation fails because of inability to compile then install the appropriate tools for your platform based on the links above and try again.

### Locating Python

When it is loaded the **reticulate** package scans the system for a compatible version of Python in the following order:

1. If specified, at the Python binary referenced by the `RETICULATE_PYTHON` environment variable.

2. The Python binary discovered on the system PATH via the `Sys.which` function.

3. At other customary locations for Python including `/usr/local/bin/python`, `/opt/local/bin/python`, etc.

You can use the `py_config` function to query for information about the specific version of Python in use as well as a list of other Python versions discovered on the system:

```{r}
library(reticulate)
py_config()
```
```

If the desired version isn't discovered automatically you should set the `RETICULATE_PYTHON` variable explicitly:

```{r}
Sys.setenv(RETICULATE_PYTHON="/usr/local/bin/python")
library(reticulate)
```

## Importing Modules

The `import` function can be used to import any Python module. For example:

```{r}
difflib <- import("difflib")
difflib$ndiff(foo, bar)

filecmp <- import("filecmp")
filecmp$cmp(dir1, dir2)
```

There are some special module names you should be aware of: `"__main__"` gives you access to the main module where code is executed by default; `"__builtin__"` gives you access to various built in Python functions (e.g. ) 

```{r}
main <- import("__main__")

py <- import("__builtin__")
py$print('foo')
```

The `"__main__"` module is generally useful if you have executed Python code from a file or a string and want to get access to it's results (see the section below for more details).

## Executing Code

You can execute Python code within the main module using the `py_run_file` and `py_run_string` functions. These functions both return a reference to the main Python module to access the results of the computations. For example:

```{r}
py_run_file("script.py")

main <- py_run_string("x = 10")
main$x
```

## Lists, Tuples, and Dictionaries

The automatic conversion of R types to Python types works well in most cases, but occasionally you will need to be more explicit on the R side to provide Python the type(s) it expects.

For example, if a Python API requires a list and you pass a single element vector it will be converted to a Python scalar. To overcome this simply use the R `list` function explicitly:

```r
foo$bar(indexes = list(42L))
```

Similarly, a Python API might require a `tuple` rather than a list. In that case you can use the `tuple` function:

```r
tuple("a", "b", "c")
```

R named lists are converted to Python dictionaries however you can also explicitly create a Python dictionary using the `dict` function:

```r
dict(foo = "bar", index = 42L)
```

This might be useful if you need to pass a dictionary that uses a more complex object (as opposed to a string) as it's key.


## With Contexts

The R `with` generic function can be used to interact with Python context manager objects (in Python you use the `with` keyword to do the same). For example:

```{r}
py <- import("__builtin__")
with(py$open("output.txt", "w") %as% file, {
  file$write("Hello, there!")
})
```

This example opens a file and ensures that it is automatically closed at the end of the execution context. Note the use of the `%as%` operator to alias the object created by the context manager.

## Iterators

If a Python API returns an [iterator or generator](http://anandology.com/python-practice-book/iterators.html) you can interact with it using the `iterate` function. The `iterate` function can be used to apply an R function to each item yielded by the iterator:

```{r}
iterate(iter, print)
```

If you don't pass a function to `iterate` the results will be collected into an R vector:

```{r}
results <- iterate(iter)
```

## Advanced Functions

There are several more advanced functions available that are useful principally when creating high level R interfaces for Python libraries: 

| Function  | Description | 
|----------------|------------------------------------------------------------------|
| py_has_attr | Check if an object has a specified attribute.  | 
| py_get_attr | Get an attribute of a specified object.  | 
| py_call | Call a Python callable object with the specified arguments.  | 
| py_capture_stdout | Capture all standard output for the specified expression and return it as an R character vector. | 
| py_suppress_warnings | Execute the specified expression, suppressing the display Python warnings.   | 



