---
title: "R Interface to Python"
output: 
  rmarkdown::html_vignette:
    readme: true
vignette: >
  %\VignetteIndexEntry{R interface to Python}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Overview

The **reticulate** package provides an R interface to Python modules, classes, and functions. For example, this code imports the Python `os` module and calls some functions within it:

```{r}
library(reticulate)
os <- import("os")
os$chdir("tests")
os$getcwd()
```

Functions and other data within Python modules and classes can be accessed via the `$` operator (analogous to the way you would interact with an R list, environment, or reference class).

When calling into Python, R data types are automatically converted to their equivalent Python types. When values are returned from Python to R they are converted back to R types. Types are converted as follows:

| R  | Python | Examples |
|---|---|---|
| Single-element vector   | Scalar |  `1`, `1L`, `TRUE`, `"foo"` |
| Multi-element vector | List  |  `c(1.0, 2.0, 3.0)`, `c(1L, 2L, 3L)` |
| List of multiple types  | Tuple  |  `list(1L, TRUE, "foo")`
| Named list | Dict  |  `list(a = 1L, b = 2.0)`, `dict(x = x_data)`
| Matrix/Array | NumPy ndarray  | `matrix(c(1,2,3,4), nrow = 2, ncol = 2)`
| Function | Python function | `function(x) x + 1`
| NULL, TRUE, FALSE  | None, True, False  |  `NULL`, `TRUE`, `FALSE`

If a Python object of a custom class is returned then an R reference to that object is returned. You can call methods and access properties of the object just as if it was an instance of an R reference class.

The **reticulate** package is compatible with all versions of Python >= 2.7 and in addition requires NumPy >= 1.11.

## Installation

You can install from GitHub as follows:

```{r}
devtools::install_github("rstudio/reticulate")
```

Note that the package includes native C/C++ code so it's installation requires [R Tools](https://cran.r-project.org/bin/windows/Rtools/) on Windows and [Command Line Tools](http://osxdaily.com/2014/02/12/install-command-line-tools-mac-os-x/) on OS X. If the package installation fails because of inability to compile then install the appropriate tools for your platform based on the links above and try again.

### Locating Python

If the version of Python you want to use is located on the system `PATH` then it will be automatically discovered (via `Sys.which`) and used. 

Alternatively, you can use one of the following functions to specify alternate versions of Python:

| Function  | Description | 
|----------------|------------------------------------------------------------------|
| use_python | Specify the path a specific Python binary. | 
| use_virtualenv | Specify the directory containing a Python virtualenv. | 
| use_condaenv | Specify the name of a Conda environment. | 

For example:

```{r}
library(reticulate)
use_python("/usr/local/bin/python")
use_virtualenv("~/myenv")
use_condaenv("myenv")
```

Note that **reticulate** requires NumPy >= 1.11 so versions of Python that don't satisfy this requirement will not be used.

Also note that the `use` functions are by default considered only hints as to where to find Python (i.e. they don't produce errors if the specified version doesn't exist). You can add the `required` parameter to ensure that the specified version of Python actually exists:

```{r}
use_virtualenv("~/myenv", required = TRUE)
```

The order in which versions of Python will be discovered and used is as follows:

1.  If specified, at the locations referenced by calls to `use_python`, `use_virtualenv`, and `use_condaenv`.

2.  If specified, at the location referenced by the `RETICULATE_PYTHON` environment variable.

3.  At the location of the Python binary discovered on the system `PATH` (via the `Sys.which` function).

4.  At other customary locations for Python including `/usr/local/bin/python`, `/opt/local/bin/python`, etc.

The scanning for and binding to a version of Python typically occurs at the time of the first call to `import` within an R session. As a result, priority will be given to versions of Python that include the module specified within the call to `import` (i.e. versions that don't include it will be skipped).

You can use the `py_config` function to query for information about the specific version of Python in use as well as a list of other Python versions discovered on the system:

```{r}
py_config()
```


## Importing Modules

The `import` function can be used to import any Python module. For example:

```{r}
difflib <- import("difflib")
difflib$ndiff(foo, bar)

filecmp <- import("filecmp")
filecmp$cmp(dir1, dir2)
```

There are some special module names you should be aware of: `"__main__"` gives you access to the main module where code is executed by default; and `"__builtin__"` gives you access to various built in Python functions. For example:

```{r}
main <- import("__main__")

py <- import("__builtin__")
py$print('foo')
```

The `"__main__"` module is generally useful if you have executed Python code from a file or string and want to get access to it's results (see the section below for more details).

## Executing Code

You can execute Python code within the main module using the `py_run_file` and `py_run_string` functions. These functions both return a reference to the main Python module so you can access the results of their execution. For example:

```{r}
py_run_file("script.py")

main <- py_run_string("x = 10")
main$x
```

## Lists, Tuples, and Dictionaries

The automatic conversion of R types to Python types works well in most cases, but occasionally you will need to be more explicit on the R side to provide Python the type it expects.

For example, if a Python API requires a list and you pass a single element R vector it will be converted to a Python scalar. To overcome this simply use the R `list` function explicitly:

```r
foo$bar(indexes = list(42L))
```

Similarly, a Python API might require a `tuple` rather than a list. In that case you can use the `tuple` function:

```r
tuple("a", "b", "c")
```

R named lists are converted to Python dictionaries however you can also explicitly create a Python dictionary using the `dict` function:

```r
dict(foo = "bar", index = 42L)
```

This might be useful if you need to pass a dictionary that uses a more complex object (as opposed to a string) as it's key.


## With Contexts

The R `with` generic function can be used to interact with Python context manager objects (in Python you use the `with` keyword to do the same). For example:

```{r}
py <- import("__builtin__")
with(py$open("output.txt", "w") %as% file, {
  file$write("Hello, there!")
})
```

This example opens a file and ensures that it is automatically closed at the end of the with block. Note the use of the `%as%` operator to alias the object created by the context manager.

## Iterators

If a Python API returns an [iterator or generator](http://anandology.com/python-practice-book/iterators.html) you can interact with it using the `iterate` function. The `iterate` function can be used to apply an R function to each item yielded by the iterator:

```{r}
iterate(iter, print)
```

If you don't pass a function to `iterate` the results will be collected into an R vector:

```{r}
results <- iterate(iter)
```

Note that the `Iterators` will be drained of their values by `iterate()`:

```{r}
a <- iterate(iter) # results are not empty
b <- iterate(iter) # results are empty since items have already been drained
```


## Callable Objects

In addition to acessing their methods and properites, some Python objects are also callable (meaning they can be invoked with parameters just like an ordinary function). Callable Python objects are returned to R as objects rather than functions, however, you can still execute the callable function via the `$call()` method, for example:

```{r}
# get a callable object
parser <- spacy$English()

# call the object as a function
parser$call(spacy)
```

## Advanced Functions

There are several more advanced functions available that are useful principally when creating high level R interfaces for Python libraries: 

| Function  | Description | 
|----------------|------------------------------------------------------------------|
| py_config | Get information on the location and version of Python in use.  | 
| py_available | Check whether a Python interface is available on this system. | 
| py_has_attr | Check if an object has a specified attribute.  | 
| py_get_attr | Get an attribute of a Python object.  | 
| py_call | Call a Python callable object with the specified arguments.  | 
| py_capture_output | Capture Python output for the specified expression and return it as an R character vector. | 
| py_suppress_warnings | Execute the specified expression, suppressing the display Python warnings.   | 
| py_unicode | Convert a string to a Python unicode object. |
| py_str | Get the string representation of Python object. |
| py_xptr_str | Evaluate an expression that prints a string with a check for a null externalptr. |
| py_is_null_xptr | Check whether a Python object is a null externalptr. |



