---
title: "Using reticulate in an R Package"
output: 
  rmarkdown::html_vignette: 
    readme: true
vignette: >
  %\VignetteIndexEntry{Using reticulate in an R Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Checking and Testing on CRAN

If you use **reticulate** in another R package you need to account for the fact that when your package is submitted to CRAN, the CRAN test servers may not have Python, NumPy, or whatever other Python modules you are wrapping in your package. If you don't do this then your package may fail to load and/or pass it's tests when run on CRAN.

There are two things you should do to ensure your package is well behaved on CRAN:

1. When importing Python modules for use inside your package you should use the `delay_load` option to ensure that the module (and Python) is loaded only on it's first use. For example:

    ```{r}
    # python 'foo' module I want to use in my package
    foo <- NULL
    
    .onLoad <- function(libname, pkgname) {
      # delay load foo module (will only be loaded when accessed via $)
      foo <<- import("foo", delay_load = TRUE)
    }
    ```

2. When writing tests, check to see if your module is available and if it isn't then skip the test. For example, if you are using the **testthat** package, you might do this:

    ```{r}
    # helper function to skip tests if we don't have the 'foo' module
    skip_if_no_foo <- function() {
      have_foo <- py_module_available("foo")
      if (!have_foo)
        skip("foo not available for testing")
    }
    
    # then call this function from all of your tests
    test_that("Things work as expected", {
      skip_if_no_foo()
      # test code here...
    })
    ```

## S3 Methods

Python objects exposed by **reticulate** carry their Python classes into R, so it's possible to write S3 methods to customize e.g. the `str` or `print` behavior for a given class (note that it's not typically necessary that you do this since the default `str` and `print` methods call `PyObject_Str`, which typically provides an acceptable default behavior).

If you do decide to implement custom S3 methods for a Python class it's important to keep in mind that when an R session ends the connection to Python objects is lost, so when the .RData saved from one R session is restored in a subsequent R session the Python objects are effectively lost (technically they become `NULL` R `externalptr` objects).

This means that you should always use the `py_is_null_xptr` function on objects before interacting with them in an S3 method. For example:

```{r}
#' @export
summary.MyPythonClass <- function(object, ...) {
  if (py_is_null_xptr(object))
    stop("Object is NULL")
  else
    # interact with the object to generate the summary
}
```

There are a couple of shortcut methods you can use to make this more straightforward. The `py_validate_xptr` function will do the check and automatically throw an error if it fails. So the example above could be re-written as:

```{r}
#' @export
summary.MyPythonClass <- function(object, ...) {
  py_validate_xptr(object)
  # interact with the object to generate the summary
}
```

Finally, the **reticulate** package exports a `py_str` generic method which is called from the `str` method only after doing appropriate validation (if the object is NULL then `<pointer: 0x0>` is returned). You can implement the `py_str` method as follows:

```{r}
#' @importFrom reticulate py_str
#' @export 
py_str.MyPythonClass <- function(object, ...) {
  # interact with the object to generate the string
}
```

So in short, to provide custom `str` and `print` methods, just implement `py_str`. For other S3 methods be sure to call either `py_validate_xptr` or `py_is_null_xptr` before interacting with the object.





