---
title: "Primer on Python for R Users"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Primer on Python for R Users}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
library(reticulate)

# this vignette requires python 3.8 or newer
eval <- tryCatch({
  config <- py_config()
  numeric_version(config$version) >= "3.8" && py_numpy_available()
}, error = function(e) FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = eval
)
```

```{r}
library(reticulate)
```

## Primer on Python for R users

You may find yourself wanting to read and understand some Python, or
even port some Python to R. This guide is designed to enable you to do
these tasks as quickly as possible. As you'll see, R and Python are
similar enough that this is possible without necessarily learning all of
Python. We start with the basics of container types and work up to the
mechanics of classes, dunders, the iterator protocol, the context
protocol, and more!

### Whitespace

Whitespace matters in Python. In R, expressions are grouped into a code
block with `{}`. In Python, that is done by making the expressions share
an indentation level. For example, an expression with an R code block
might be:

```{r}
if (TRUE) {
  cat("This is one expression. \n")
  cat("This is another expression. \n")
}
```

The equivalent in Python:

```{python}
if True:
  print("This is one expression.")
  print("This is another expression.")
```

Python accepts tabs or spaces as the indentation spacer, but the rules
get tricky when they're mixed. Most style guides suggest (and IDE's
default to) using spaces only.

### Container Types

In R, the `list()` is a container you can use to organize R objects. R's
`list()` is feature packed, and there is no single direct equivalent in
Python that supports all the same features. Instead there are (at least)
4 different Python container types you need to be aware of: lists,
dictionaries, tuples, and sets.

#### Lists

Python lists are typically created using bare brackets `[]`. The Python
built-in `list()` function is more of a coercion function, closer in
spirit to R's `as.list()`. The most important thing to know about Python
lists is that they are modified in place. Note in the example below that
`y` reflects the changes made to `x`, because the underlying list object
which both symbols point to is modified in place.

```{python}
x = [1, 2, 3]
y = x    # `y` and `x` now refer to the same list!
x.append(4)
print("x is", x)
print("y is", y)
```

One Python idiom that might be concerning to R users is that of growing
lists through the `append()` method. Growing lists in R is typically
slow and best avoided. But because Python's list are modified in place
(and a full copy of the list is avoided when appending items), it is
efficient to grow Python lists in place.

Some syntactic sugar around Python lists you might encounter is the
usage of `+` and `*` with lists. These are concatenation and replication
operators, akin to R's `c()` and `rep()`.

```{python}
x = [1]
x
x + x
x * 3
```

You can index into lists with integers using trailing `[]`, but note
that indexing is 0-based.

```{python}
x = [1, 2, 3]
x[0]
x[1]
x[2]

try:
  x[3]
except Exception as e:
  print(e)
```

When indexing, negative numbers count from the end of the container.

```{python}
x = [1, 2, 3]
x[-1]
x[-2]
x[-3]
```

You can slice ranges of lists using the `:` inside brackets. Note that
the slice syntax is ***not*** inclusive of the end of the slice range.
You can optionally also specify a stride.

```{python}
x = [1, 2, 3, 4, 5, 6]
x[0:2] # get items at index positions 0, 1
x[1:]  # get items from index position 1 to the end
x[:-2] # get items from beginning up to the 2nd to last.
x[:]   # get all the items (idiom used to copy the list so as not to modify in place)
x[::2] # get all the items, with a stride of 2
x[1::2] # get all the items from index 1 to the end, with a stride of 2
```

#### Tuples

Tuples behave like lists, except they are not mutable, and they don't
have the same modify-in-place methods like `append()`. They are
typically constructed using bare `()`, but parentheses are not strictly
required, and you may see an implicit tuple being defined just from a
comma separated series of expressions. Because parentheses can also be
used to specify order of operations in expressions like `(x + 3) * 4`, a
special syntax is required to define tuples of length 1: a trailing
comma. Tuples are most commonly encountered in functions that take a
variable number of arguments.

```{python}
x = (1, 2) # tuple of length 2
type(x)
len(x)
x

x = (1,) # tuple of length 1
type(x)
len(x)
x

x = () # tuple of length 0
print(f"{type(x) = }; {len(x) = }; {x = }")
# example of an interpolated string literals

x = 1, 2 # also a tuple
type(x)
len(x)

x = 1, # beware a single trailing comma! This is a tuple!
type(x)
len(x)
```

##### Packing and Unpacking

Tuples are the container that powers the *packing* and *unpacking*
semantics in Python. Python provides the convenience of allowing you to
assign multiple symbols in one expression. This is called *unpacking*.

For example:

```{python}
x = (1, 2, 3)
a, b, c = x
a
b
c
```

(You can access similar unpacking behavior from R using
`` zeallot::`%<-%` ``).

Tuple unpacking can occur in a variety of contexts, such as iteration:

```{python}
xx = (("a", 1),
      ("b", 2))
for x1, x2 in xx:
  print("x1 = ", x1)
  print("x2 = ", x2)
```

If you attempt to unpack a container to the wrong number of symbols,
Python raises an error:

```{python, error = TRUE}
x = (1, 2, 3)
a, b, c = x # success
a, b = x    # error, x has too many values to unpack
a, b, c, d = x # error, x has not enough values to unpack
```

It is possible to unpack a variable number of arguments, using `*` as a
prefix to a symbol. (You'll see the `*` prefix again when we talk about
functions)

```{python}
x = (1, 2, 3)
a, *the_rest = x
a
the_rest
```

You can also unpack nested structures:

```{python}
x = ((1, 2), (3, 4))
(a, b), (c, d) = x
```

#### Dictionaries

Dictionaries are most similar to R environments. They are a container
where you can retrieve items by name, though in Python the name (called
a *key* in Python's parlance) does not need to be a string like in R. It
can be any Python object with a `hash()` method (meaning, it can be
almost any Python object). They can be created using syntax like
`{key: value}`. Like Python lists, they are modified in place. Note that
`r_to_py()` converts R named lists to dictionaries.

```{python}
d = {"key1": 1,
     "key2": 2}
d2 = d
d
d["key1"]
d["key3"] = 3
d2 # modified in place!
```

Like R environments (and unlike R's named lists), you cannot index into
a dictionary with an integer to get an item at a specific index
position. Dictionaries are *unordered* containers. (However---beginning
with Python 3.7, dictionaries do preserve the item insertion order).

```{python, error = TRUE}
d = {"key1": 1, "key2": 2}
d[1] # error
```

A container that closest matches the semantics of R's named list is the
[`OrderedDict`](https://docs.python.org/3/library/collections.html#collections.OrderedDict),
but that's relatively uncommon in Python code so we don't cover it
further.

#### Sets

Sets are a container that can be used to efficiently track unique items
or deduplicate lists. They are constructed using `{val1, val2}` (like a
dictionary, but without `:`). Think of them as dictionary where you only
use the keys. Sets have many efficient methods for membership
operations, like `intersection()`, `issubset()`, `union()` and so on.

```{python}
s = {1, 2, 3}
type(s)
s

s.add(1)
s
```

### Iteration with `for`

The `for` statement in Python can be used to iterate over any kind of
container.

```{python}
for x in [1, 2, 3]:
  print(x)
```

R has a relatively limited set of objects that can be passed to `for`.
Python by comparison, provides an iterator protocol interface, which
means that authors can define custom objects, with custom behavior that
is invoked by `for`. (We'll have an example for how to define a custom
iterable when we get to classes). You may want to use a Python iterable
from R using reticulate, so it's helpful to peel back the syntactic
sugar a little to show what the `for` statement is doing in Python, and
how you can step through it manually.

There are two things that happen: first, an iterator is constructed from
the supplied object. Then, the new iterator object is repeatedly called
with `next()` until it is exhausted.

```{python, error = TRUE}
l = [1, 2, 3]
it = iter(l) # create an iterator object
it

# call `next` on the iterator until it is exhausted:
next(it)
next(it)
next(it)
next(it)
```

In R, you can use reticulate to step through an iterator the same way.

```{r}
library(reticulate)
l <- r_to_py(list(1, 2, 3))
it <- as_iterator(l)

iter_next(it)
iter_next(it)
iter_next(it)
iter_next(it, completed = "StopIteration")
```

Iterating over dictionaries first requires understanding if you are
iterating over the keys, values, or both. Dictionaries have methods that
allow you to specify which.

```{python}
d = {"key1": 1, "key2": 2}
for key in d:
  print(key)

for value in d.values():
  print(value)

for key, value in d.items():
  print(key, ":", value)
```

#### Comprehensions

Comprehensions are special syntax that allow you to construct a
container like a list or a dict, while also executing a small operation
or single expression on each element. You can think of it as special
syntax for R's `lapply`.

For example:

```{python}
x = [1, 2, 3]

# a list comprehension built from x, where you add 100 to each element
l = [element + 100 for element in x]
l

# a dict comprehension built from x, where the key is a string.
# Python's str() is like R's as.character()
d = {str(element) : element + 100
     for element in x}
d
```

### Defining Functions with `def`

Python functions are defined with the `def` statement. The syntax for
specifying function arguments and default values is very similar to R.

```{python}
def my_function(name = "World"):
  print("Hello", name)

my_function()
my_function("Friend")
```

The equivalent R snippet would be

```{r}
my_function <- function(name = "World") {
  cat("Hello", name, "\n")
}

my_function()
my_function("Friend")
```

Unlike R functions, the last value in a function is not automatically
returned. Python requires an explicit return statement.

```{python}
def fn():
  1
print(fn())

def fn():
  return 1
print(fn())
```

(Note for advanced R users: Python has no equivalent of R's argument
"promises". Function argument default values are evaluated once, when
the function is constructed. This can be surprising if you define a
Python function with a mutable object as a default argument value, like
a Python list!)

```{python}
def my_func(x = []):
  x.append("was called")
  print(x)

my_func()
my_func()
my_func()
```

You can also define Python functions that take a variable number of
arguments, similar to `...` in R. A notable difference is that R's `...`
makes no distinction between named and unnamed arguments, but Python
does. In Python, prefixing a single `*` captures unnamed arguments, and
two `**` signifies that *keyword* arguments are captured.

```{python}
def my_func(*args, **kwargs):
  print("args = ", args) # args is a tuple
  print("kwargs = ", kwargs) # kwargs is a dictionary

my_func(1, 2, 3, a = 4, b = 5, c = 6)
```

Whereas the `*` and `**` in a function definition signature *pack*
arguments, in a function call they *unpack* arguments. Unpacking
arguments in a function call is equivalent to using `do.call()` in R.

```{python}
def my_func(a, b, c):
  print(a, b, c)

args = (1, 2, 3)
my_func(*args)

kwargs = {"a": 1, "b": 2, "c": 3}
my_func(**kwargs)
```

### Defining Classes with `class`

One could argue that in R, the preeminent unit of composition for code
is the `function`, and in Python, it's the `class`. You can be a very
productive R user and never use R6, reference classes, or similar R
equivalents to the object-oriented style of Python `class`'s.

In Python, however, understanding the basics of how `class` objects work
is requisite knowledge, because `class`'s are how you organize and find
methods in Python. (In contrast to R's approach, where methods are found
by dispatching from a generic). Fortunately, the basics of `class`'s are
accessible.

Don't be intimidated if this is your first exposure to object oriented
programming. We'll start by building up a simple Python class for
demonstration purposes.

```{python}
class MyClass:
  pass # `pass` means do nothing.

MyClass
type(MyClass)

instance = MyClass()
instance
type(instance)
```

Like the `def` statement, the `class` statement binds a new callable
symbol, `MyClass`. First note the strong naming convention, classes are
typically `CamelCase`, and functions are typically `snake_case`. After
defining `MyClass`, you can interact with it, and see that it has type
`'type'`. Calling `MyClass()` creates a new object **instance** of the
class, which has type `'MyClass'` (ignore the `__main__.` prefix for
now). The instance prints with its memory address, which is a strong
hint that it's common to be managing many instances of a class, and that
the instance is mutable (modified-in-place by default).

In the first example, we defined an empty `class`, but when we inspect
it we see that it already comes with a bunch of attributes (`dir()` in
Python is equivalent to `names()` in R):

```{python}
dir(MyClass)
```

#### What are all the underscores?

Python typically indicates that something is special by wrapping the
name in double underscores. A special double-underscore-wrapped token is
commonly called a "dunder". "Special" is not a technical term, it just
means that the token invokes a Python language feature. Some dunder
tokens are merely ways code authors can plug into specific syntactic
sugars, others are values provided by the interpreter that would be
otherwise hard to acquire, yet others are for extending language
interfaces (e.g., the iteration protocol), and finally, a small handful
of dunders are truly complicated to understand. Fortunately, as an R
user looking to use some Python features through reticulate, you only
need to know about a few easy-to-understand dunders.

The most common dunder method you'll encounter when reading Python code
is `__init__()`. This is a function that is called when the class
constructor is called, that is, when a class is **instantiated**. It is
meant to initialize the new class instance. (In very sophisticated code
bases, you may also encounter classes where `__new__` is also defined,
this is called before `__init__`).

```{python}
class MyClass:

  print("MyClass's definition body is being evaluated")

  def __init__(self):
    print(self, "is initializing")

print("MyClass is finished being created")

instance = MyClass()
print(instance)

instance2 = MyClass()
print(instance2)
```

A few things to note:

-   the `class` statement takes a code block that is defined by a common
    indentation level. The code block has the same exact semantics as
    any other expression that takes a code block, like `if` and `def`.
    The body of the class is evaluated only **once**, when the class
    constructor is first being created. Beware that any objects defined
    here are shared by all instances of the class!

-   `__init__` is just a normal function, defined with `def` like any
    other function. Except it's inside the class body.

-   `__init__` take an argument: `self`. `self` is the class instance
    being initialized (note the identical memory address between `self`
    and `instance`). Also note that we didn't provide `self` when call
    `MyClass()` to create the class instance, `self` was spliced into
    the function call by the interpreter.

-   `__init__` is called each time a new instance is created.

Functions defined inside a `class` code block are called *methods*, and
the important thing to know about methods is that each time they are
called from a class instance, the instance is spliced into the function
call as the first argument. This applies to all functions defined in a
class, including dunders. (The sole exception is if the function is
decorated with something like `@classmethod` or `@staticmethod`).

```{python, error = TRUE}
class MyClass:
  def a_method(self):
    print("MyClass.a_method() was called with", self)

instance = MyClass()
instance.a_method()

MyClass.a_method()     # error, missing required argument `self`
MyClass.a_method(instance) # identical to instance.a_method()
```

Other dunder's worth knowing about are:

-   `__getitem__`: the function invoked when subsetting an instance with
    `[` (Equivalent to defining a `[` S3 method in R.

-   `__getattr__`: the function invoked when subsetting with `.`
    (Equivalent to defining a `$` S3 method in R.

-   `__iter__` and `__next__`: functions invoked by `for`.

-   `__call__`: invoked when a class instance is called like a function
    (e.g., `instance()`).

-   `__bool__`: invoked by `if` and `while` (equivalent to
    `as.logical()` in R, but returning only a scalar, not a vector).

-   `__repr__`, `__str__`, functions invoked for formatting and pretty
    printing (akin to `format()`, `dput()`, and `print()` methods in R).

-   `__enter__` and `__exit__`: functions invoked by `with`.

-   Many [built-in](https://docs.python.org/3/library/functions.html)
    Python functions are just sugar for invoking the dunder. For
    example: calling `repr(x)` is identical to `x.__repr__()`. Other
    builtins that are just sugar for invoking the dunder are `next()`,
    `iter()`, `str()`, `list()`, `dict()`, `bool()`, `dir()`, `hash()`
    and more!

#### Iterators, revisited

Now that we have the basics of `class`, it's time to revisit iterators.
First, some terminology:

**iterable**: something that can be iterated over. Concretely, a class
that defines an `__iter__` method, whose job is to return an *iterator*.

**iterator**: something that iterates. Concretely, a class that defines
a `__next__` method, whose job is to return the next element each time
it is called, and then raises a `StopIteration` exception once it's
exhausted.

It's common to see classes that are both iterables and iterators, where
the `__iter__` method is just a stub that returns `self`.

Here is a custom iterable / iterator implementation of Python's `range`
(similar to `seq` in R)

```{python, error = TRUE}
class MyRange:
  def __init__(self, start, end):
    self.start = start
    self.end = end

  def __iter__(self):
    # reset our counter.
    self._index = self.start - 1
    return self

  def __next__(self):
    if self._index < self.end:
      self._index += 1 # increment
      return self._index
    else:
      raise StopIteration

for x in MyRange(1, 3):
  print(x)

# doing what `for` does, but manually
r = MyRange(1, 3)
it = iter(r)
next(it)
next(it)
next(it)
next(it)
```

### Defining Generators with `yield`.

Generators are special Python functions that contain one or more `yield`
statements. As soon as `yield` is included in a code block passed to
`def`, the semantics change substantially. You're no longer defining a
mere function, but a generator constructor! In turn, calling a generator
constructor creates a generator object, which is just another type of
iterator.

Here is an example:

```{python, error = TRUE}
def my_generator_constructor():
  yield 1
  yield 2
  yield 3

# At first glance it presents like a regular function
my_generator_constructor
type(my_generator_constructor)

# But calling it returns something special, a 'generator object'
my_generator = my_generator_constructor()
my_generator
type(my_generator)

# The generator object is both an iterable and an iterator
# it's __iter__ method is just a stub that returns `self`
iter(my_generator) == my_generator == my_generator.__iter__()

# step through it like any other iterator
next(my_generator)
my_generator.__next__() # next() is just sugar for calling the dunder
next(my_generator)
next(my_generator)
```

Encountering `yield` is like hitting the pause button on a functions
execution, it preserves the state of everything in the function body and
returns control to whatever is iterating over the generator object.
Calling `next()` on the generator object resumes execution of the
function body until the next `yield` is encountered, or the function
finishes.

### Iteration closing remarks

Iteration is deeply baked into the Python language, and R users may be
surprised by how things in Python are iterable, iterators, or powered by
the iterator protocol under the hood. For example, the built-in `map()`
(equivalent to R's `lapply()`) yields an iterator, not a list.
Similarly, a tuple comprehension like `(elem for elem in x)` produces an
iterator. Most features dealing with files are iterators, and so on.

Any time you find an iterator inconvenient, you can materialize all the
elements into a list using the Python built-in `list()`, or
`reticulate::iterate()` in R. Also, if you like the readability of
`for`, you can utilize similar semantics to Python's `for` using
`coro::loop()`.

### `import` and Modules

In R, authors can bundle their code into shareable extensions called R
packages, and R users can access objects from R packages via `library()`
or `::`. In Python, authors bundle code into *modules*, and users access
modules using `import`. Consider the line:

```{python}
import numpy
```

This statement has Python go out to the file system, find an installed
Python module named 'numpy', load it (commonly meaning: evaluate its
`__init__.py` file and construct a `module` type), and bind it to the
symbol `numpy`.

The closest equivalent to this in R might be:

```{r, eval = FALSE}
dplyr <- loadNamespace("dplyr")
```

#### Where are modules found?

In Python, the file system locations where modules are searched can be
accessed (and modified) from the list found at `sys.path`. This is
Python's equivalent to R's `.libPaths()`. `sys.path` will typically
contain paths to the current working directory, the Python installation
which contains the built-in standard library, administrator installed
modules, user installed modules, values from environment variables like
`PYTHONPATH`, and any modifications made directly to `sys.path` by other
code in the current Python session (though this is relatively uncommon
in practice).

```{python}
import sys
sys.path
```

You can inspect where a module was loaded from by accessing the dunder
`__path__` or `__file__` (especially useful when troubleshooting
installation issues):

```{python}
import os
os.__file__
numpy.__path__
```

Once a module is loaded, you can access symbols from the module using
`.` (equivalent to `::`, or maybe `$.environment`, in R).

```{python}
numpy.abs(-1)
```

There is also special syntax for specifying the symbol a module is bound
to upon import, and for importing only some specific symbols.

```{python}
import numpy        # import
import numpy as np  # import and bind to a custom symbol `np`
np is numpy         # test for identicalness, similar to identical(np, numpy)

from numpy import abs # import only `numpy.abs`, bind it to `abs`
abs is numpy.abs

from numpy import abs as abs2 # import only `numpy.abs`, bind it to `abs2`
abs2 is numpy.abs
```

If you're looking for the Python equivalent of R's `library()`, which
makes all of a package's exported symbols available, it might be using
`import` with a `*` wildcard, though it's relatively uncommon to do so.
The `*` wildcard will expand to include all the symbols in module, or
all the symbols listed in `__all__`, if it is defined.

```{python}
from numpy import *
```

Python doesn't make a distinction like R does between package exported
and internal symbols. In Python, all module symbols are equal, though
there is the naming convention that intended-to-be-internal symbols are
prefixed with a single leading underscore. (Two leading underscores
invoke an advanced language feature called "name mangling", which is
outside the scope of this introduction).

### Integers and Floats

R users generally don't need to be aware of the difference between
integers and floating point numbers, but that's not the case in Python.
If this is your first exposure to numeric data types, here are the
essentials:

-   integer types can only represent whole numbers like `1` or `2`, not
    floating point numbers like `1.2`.

-   floating-point types can represent any number, but with some degree
    of imprecision.

In R, writing a bare literal number like `12` produces a floating point
type, whereas in Python, it produces an integer. You can produce an
integer literal in R by appending an `L`, as in `12L`. Many Python
functions expect integers, and will error when provided a float.

For example, say we have a Python function that expects an integer:

```{python}
def a_strict_Python_function(x):
  assert isinstance(x, int), "x is not an int"
  print("Yay! x was an int")
```

When calling it from R, you must be sure to call it with an integer:

```{r, error = TRUE}
library(reticulate)
py$a_strict_Python_function(3)             # error
py$a_strict_Python_function(3L)            # success
py$a_strict_Python_function(as.integer(3)) # success
```

### What about R vectors?

R is a language designed for numerical computing first. Numeric vector
data types are baked deep into the R language, to the point that the
language doesn't even distinguish scalars from vectors. By comparison,
numerical computing capabilities in Python are generally provided by
third party packages (*modules*, in Python parlance).

In Python, the `numpy` module is most commonly used to handle contiguous
arrays of data. The closest equivalent to an R numeric vector is a numpy
array, or sometimes, a list of scalar numbers (some Pythonistas might
argue for `array.array()` here, but that's so rarely encountered in
actual Python code we don't mention it further).

Teaching the NumPy interface is beyond the scope of this primer, but
it's worth pointing out some potential tripping hazards for users
accustomed to R arrays:

-   When indexing into multidimensional numpy arrays, trailing
    dimensions can be omitted and are implicitly treated as missing. The
    consequence is that iterating over arrays means iterating over the
    first dimension. For example, this iterates over the rows of a
    matrix.

```{python}
import numpy as np
m = np.arange(12).reshape((3,4))
m
m[0, :] # first row
m[0]    # also first row

for row in m:
  print(row)
```

-   Many numpy operations modify the array in place! This is surprising
    to R users, who are used to the convenience and safety of R's
    copy-on-modify semantics. Unfortunately, there is no simple scheme
    or naming convention you can rely on to quickly determine if a
    particular method modifies in-place or creates a new array copy. The
    only reliable way is to consult the
    [documentation](https://numpy.org/doc/stable/reference/index.html#reference),
    and conduct small experiments at the `reticulate::repl_python()`.

### Decorators

Decorators are just functions that take a function as an argument, and
then typically returns another function. Any function can be invoked as
a decorator with the `@` syntax, which is just sugar for this simple
action:

```{python}
def my_decorator(func):
  func.x = "a decorator modified this function by adding an attribute `x`"
  return func

def my_function(): pass
my_function = my_decorator(my_function)

# @ is just fancy syntax for the above two lines
@my_decorator
def my_function(): pass
```

One decorator you might encounter frequently is:

-   `@property`, which automatically calls a class method when the
    attribute is accessed (similar to `makeActiveBinding()` in R).

### `with` and context management

Any object that defines `__enter__` and `__exit__` methods implements
the "context" protocol, and can be passed to `with`. For example, here
is a custom implementation of a context manager that temporarily changes
the current working directory (equivalent to R's `withr::with_dir()`)

```{python}
from os import getcwd, chdir

class wd_context:
  def __init__(self, wd):
    self.new_wd = wd

  def __enter__(self):
    self.original_wd = getcwd()
    chdir(self.new_wd)

  def __exit__(self, *args):
    # __exit__ takes some additional argument that are commonly ignored
    chdir(self.original_wd)


getcwd()
with wd_context(".."):
  print("in the context, wd is:", getcwd())
getcwd()
```

### Learning More

Hopefully, this short primer to Python has provided a good foundation
for confidently reading Python documentation and code, and using Python
modules from R via reticulate. Of course, there is much, much more to
learn about Python. Googling questions about Python reliably brings up
pages of results, but not always sorted in order of most useful. Blog
posts and tutorials targeting beginners can be valuable, but remember
that Python's official documentation is generally excellent, and it
should be your first destination when you have questions.

<https://docs.Python.org/3/>

<https://docs.Python.org/3/library/index.html>

To learn Python more fully, the built-in official tutorial is also
excellent and comprehensive (but does require a time commitment to get
value out of it) <https://docs.Python.org/3/tutorial/index.html>

Finally, don't forget to solidify your understanding by conducting small
experiments at the `reticulate::repl_python()`.

Thank you for reading!
